\chapter*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

Having recently read A.J. Jacobs' \emph{Thanks a  Thousand} and  being of persuadable character, I am inclined to search out and thank everyone who has helped with this thesis and made my time at Oxford so extraordinary. 

This, of course, would be a terrible idea. Where would I stop? 
Alas, resigned to the fact  that I must commit a kind of thanksism  and draw  an arbitrary line between those I thank and  those I don't, I do however wish to extend  my gratitude to several  people. 

Thanks is due to Renaud, who unflinchingly allowed an all-too-enthusiastic master's student to pursue his topic of choice, and follow any path  he found interesting. This freedom was  greatly appreciated. I also owe a great  deal of thanks to Karel,  who provided insight into the topic which I would  have never  garnered  by myself, and who  was  always willing to answer my (incessant) questions---even while  on holiday. He was also a constant source  of motivation and  proofread  a draft  of this work. 

This year would not have been one tenth as exciting (albeit perhaps ten times as productive) without the St. Cross student community and the Wolfson College Boat Club. Though small in quadrangle count (it's  how you use them  that counts), St. Cross proved itself  full of friendly and talented students. More importantly, it provided lunch---by far the most  epic on campus. 
The yogurt by itself redeems the odd smell in the library that nobody wants to talk about. 
WCBC ensured that I caught many crabs, lengthened  my strokes, and perfected my hip swing. Sometimes I even rowed. While I  can no longer drink gin  without thinking of Denmark, I could not have  asked for a  better group of people  with whom to be trapped in  a boat early in the morning, being yelled at about pineapples and early squares. In the words of a large Swiss gentlemen, this  year ``was better than ..." 

My parents provide a constant source of support and encouragement for which I am  incredibly  grateful. I  would be nowhere without you both.  

Much love to all. 


